### 2014-10-10 writing/AAAI SSS '15/join.tex
Minor edits
Misc clarifications.

\subsection*{Reasoning about semantic relations}\label{sec:join}

@@

If a model is to learn the behavior of a relational logic like the natural
logic presented here over some domain from a finite amount of data, it 
must learn to deduce new relational statements from seen ones in a sound manner.
The simplest such deductions, and the only ones that can be made using relations
between single terms, involve the set of inference rules schematized in 
Figure~\ref{tab:jointable}. For instance, given that $a \natrev b$ and $b
\natrev c$, one can conclude that $a \natrev c$, by basic
set-theoretic reasoning (transitivity of $\natrev$). Similarly, from
$a \natfor b$ and $b \natneg c$, it follows that $a \natalt c$. Cells in the
table containing a dot correspond to pairs of relations for which no valid 
inference can be drawn in our logic.

@@

\paragraph{Results} 
We found that the NTN model worked best with 11-dimensional vector
representations for the 80 sets and a 90-dimensional feature vector
for the classifier, though the performance of the model was not highly
sensitive to either dimensionality parameter in any of the experiments
discussed here. This model was able to correctly label 99.4\% (94.3
macroaveraged F1) of the provably derivable test examples, and 95.4\%
(92.2 macro-F1) of the remaining test examples. The simpler NN worked
best with 11 and 75 dimensions, respectively, but was able to achieve
accuracies of only 89.0\% (80.7 macro-F1) and 62.5\% (40.7 macro-F1),
respectively. Training accuracy was 99.8\% for the NTN and 95.4\% for
the NN.

>>>>

\section{Reasoning about semantic relations}\label{sec:join}

@@

In this experiment, we test our model's ability to learn and use
the natural logic inference rules schematized in 
Figure~\ref{tab:jointable}. For instance, given that $a \natrev b$ and $b
\natrev c$, one can conclude that $a \natrev c$, by basic
set-theoretic reasoning (transitivity of $\natrev$). Similarly, from
$a \natfor b$ and $b \natneg c$, it follows that $a \natalt c$. Cells in the
table containing a dot correspond to pairs of relations for which no valid 
inference can be drawn in our logic.

@@

\paragraph{Results} 
We found that the NTN model worked best with 11-dimensional vector
representations for the 80 sets and a 90-dimensional feature vector
for the classifier, though the performance of the model was not highly
dependant on either dimensionality setting. 
This model was able to correctly label 99.4\% (94.3
macroaveraged F1) of the provably derivable test examples, and 95.4\%
(92.2 macro-F1) of the remaining test examples. The simpler NN worked
best with 11 and 75 dimensions, respectively, but was able to achieve
accuracies of only 89.0\% (80.7 macro-F1) and 62.5\% (40.7 macro-F1),
respectively. Training accuracy was 99.8\% for the NTN and 95.4\% for
the NN.

### 2014-10-10 writing/AAAI SSS '15/wordnet.tex
Misc clarifications.

Using simulated data as above is reassuring about what the models
learn and why, but we also want to know how they perform with a real
natural language vocabulary. Unfortunately, as far as we are aware,
there are no available resources labeling such a vocabulary with the
relations from Table~\ref{b-table}. However, the relations in WordNet
\cite{fellbaum2010wordnet} come close and pose the same substantive
challenges.

@@

We extract three types of relation from WordNet. \ii{Hypernym} and
\ii{hyponym} can be are represented directly in the WordNet graph
structure, and correspond closely to the $\natrev$ and $\natfor$
relations from natural logic. As in natural logic, these relations are
mirror images of one another: if \ii{dog} is a hyponym of \ii{animal}
(perhaps indirectly by way of \ii{canid}, \ii{mammal}, etc.), then
\ii{animal} is a hypernym of \ii{dog}. We also extract \ii{coordinate}
terms, which share a direct hypernym, like \ii{dalmatian}, \ii{pug},
and \ii{puppy}, which are all direct hyponyms of \ii{dog}.  Coordinate
terms tend to exclude one another, thereby providing a reasonable
approximation of exclusion. WordNet defines hypernymy and hyponymy over
sets of synonyms, rather than over individual terms, so we do not
include a \ii{synonym} or \ii{equivalent} relation, but rather
consider only one member of each set of synonyms.

To limit the size of the vocabulary without otherwise simplifying the learning problem, we extract all of the
instances of these three relations for single word nouns in WordNet that are hyponyms of the node 
\texttt{organism.n.01}. In order to balance the distribution of the classes, we slightly downsample instances 
of the \ii{coordinate} relation, yielding 36,772 relations among 3,217 terms. We randomly select 15\% of these 
relations as test data. Unlike in the previous experiment, it is not straightforward here to determine in advance
 how much training data should be required to learn an accurate model, so we performed training runs with 
 various fractions of the training portion of the 
dataset. Embeddings were fixed at 25 dimensions and were initialized randomly or using distributional 
vectors from GloVe \cite{pennington2014glove}. The feature vector produced by the comparison layer was fixed at 
80 dimensions.

@@

\begin{table}[h]
\centering\resizebox{5.5in}{!}{
  \setlength{\tabcolsep}{15pt}
  \renewcommand{\arraystretch}{1.1}
  \begin{tabular}{r r@{ }r r@{ }r r@{ }r r@{ }r r@{ }r} 
    \toprule
     Portion of & \multicolumn{4}{c}{NN} & \multicolumn{4}{c}{NTN} & \multicolumn{2}{c}{Baseline}\\
     training data  & \multicolumn{2}{c}{w/ GloVe} & \multicolumn{2}{c}{w/o GloVe} & \multicolumn{2}{c}{w/ GloVe} & \multicolumn{2}{c}{w/o GloVe} \\
    \midrule
     100\%   & 99.5 &(100.0) & 98.9 &(99.7)    & \textbf{99.6} &(100.0) & 99.2 &(100.0) & 37.9 &(--)\\
     75\%    & 98.9 &(100.0) & 37.9 &(37.0)    & 98.6 &(100.0)          & 98.4 &(100.0) & 37.9 &(--)\\
     33\%    & 96.0 &(100.0) & 37.0 &(37.0)    & 95.4 &(100.0)          & 95.3 &(100.0) & 37.9 &(--)\\ % TODO: Rerunning. Got stuck predicting the most frequent class.
     16\%    & 91.8 &(100.0) & 91.9 &(100.0)   & 92.1 &(99.7)           & 92.2 &(100.0) & 37.9 &(--)\\
    \bottomrule
  \end{tabular}}
 \caption{Test (train) accuracy figures on the WordNet data. The baseline figure is simply the frequency of the most frequent class, \ii{coordinate}.\label{b-table}}  
\end{table}

>>>>

Using simulated data as above is reassuring about what the models
learn and why, but we also want to know how they perform with a real
natural language vocabulary. Unfortunately, as far as we are aware,
there are no available resources labeling such a vocabulary with the
relations from Table~\ref{b-table}. However, the relations in WordNet
\cite{fellbaum2010wordnet} come close and pose the same substantive
challenges within a somewhat easier classification problem.

@@

We extract three types of relation from WordNet. \ii{Hypernym} and
\ii{hyponym} can be are represented directly in the WordNet graph
structure, and correspond closely to the $\natrev$ and $\natfor$
relations from natural logic. As in natural logic, these relations are
mirror images of one another: if \ii{dog} is a hyponym of \ii{animal}
(perhaps indirectly by way of \ii{canid}, \ii{mammal}, etc.), then
\ii{animal} is a hypernym of \ii{dog}. We also extract \ii{coordinate}
terms, which share a direct hypernym, like \ii{dalmatian}, \ii{pug},
and \ii{puppy}, which are all direct hyponyms of \ii{dog}.  Coordinate
terms tend to exclude one another, thereby providing a loose approximation 
of the natural logic exclusion relation $\natneg$. 
WordNet defines hypernymy and hyponymy over
sets of synonyms, rather than over individual terms, so we do not
include a \ii{synonym} or \ii{equivalent} relation, but rather
consider only one member of each set of synonyms. Word pairs which do
not fall into these three relations are not included in the dataset.

To limit the size of the vocabulary without otherwise simplifying the learning problem, we extract all of the
instances of these three relations for single word nouns in WordNet that are hyponyms of the node 
\texttt{organism.n.01}. In order to balance the distribution of the classes, we slightly downsample instances 
of the \ii{coordinate} relation, yielding a total of 36,772 relations among 3,217 terms. We report results below using crossvalidation, choosing a disjoin 10\% test sample for each of five runs. Unlike in the previous experiment,
it is not straightforward here to determine in advance
how much data should be required to train an accurate model, so we performed training runs with 
various fractions of the remaining data. Embeddings were fixed at 25 dimensions and were initialized 
randomly or using distributional vectors from GloVe \cite{pennington2014glove}. The feature vector 
produced by the comparison layer was fixed at 80 dimensions.

@@

\begin{table}[h]
\centering\resizebox{5.5in}{!}{
  \setlength{\tabcolsep}{15pt}
  \renewcommand{\arraystretch}{1.1}
  \begin{tabular}{r r@{ }r r@{ }r r@{ }r r@{ }r r@{ }r} 
    \toprule
     Portion of & \multicolumn{4}{c}{NN} & \multicolumn{4}{c}{NTN} & \multicolumn{2}{c}{Baseline}\\
     training data  & \multicolumn{2}{c}{w/ GloVe} & \multicolumn{2}{c}{w/o GloVe} & \multicolumn{2}{c}{w/ GloVe} & \multicolumn{2}{c}{w/o GloVe} \\
    \midrule
     100\%   & 99.5 &(100.0) & 98.9 &(99.7)    & \textbf{99.6} &(100.0) & 99.2 &(100.0) & 37.9 &(--)\\
     33\%    & 96.0 &(100.0) & 37.0 &(37.0)    & 95.4 &(100.0)          & 95.3 &(100.0) & 37.9 &(--)\\
     11\%    & 91.8 &(100.0) & 91.9 &(100.0)   & 92.1 &(99.7)           & 92.2 &(100.0) & 37.9 &(--)\\
     % 2\%    & 86.0 &(100.0) & ... &(100.0)   & 86.2 &(100.0)           & 76.4 &(87.3) & 37.9 &(--)\\ % Accidental run with slightly different train/test split.
    \bottomrule
  \end{tabular}}
 \caption{Test (train) accuracy figures on the WordNet data. The baseline figure is simply the frequency of the most frequent class, \ii{coordinate}.\label{b-table}}  
\end{table}
